---
title: "varroa Ploidy"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
editor_options: 
  chunk_output_type: console
---

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```

load libraries
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("tidyverse")
library("plyr")
library("dplyr")
library("ggplot2")
library("RColorBrewer")
library("data.table")
library("stringr")
library("janitor")
library("knitr")
library("kableExtra")
library("plotly") 

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
                 #    fig.width = 20,
                  #    fig.asp = 0.6,
                   #   out.width = "100%")
```

load data
```{r, warning=FALSE}
data <- read.csv("/Users/nuriteliash/Documents/GitHub/varroa_Ploidy/data/Ploidy.csv") %>%
  dplyr::mutate(Family = as.character(Family)) %>%
  dplyr::mutate(stage_mature = case_when(
    grepl("adult", Stage) ~ "Mature",
    !grepl("adult", Stage) ~ "Imature"))

# order the levels 
data$body.part <- factor(data$body.part, level=c("Body", "Anterior", "Posterior", "Legs","Hemolymph","Ovary","Testes"))

data$Stage <- factor(data$Stage, level=c("Larvae", "Protonymph", "Deuteronymph", "adult"))

data$Stage_original <- factor(data$Stage_original, level=c("Mom", "Son", "Mature","Daughter", "Deuteronymph", "Protonymph", "Larvae", "Immature"))

levels(data$Family) <- c(levels(data$Family), 0)
data$Family <- factor(data$Family, level= c("0", "1", "11","27","3","2","4", "5"))

p_family_adults = data %>% dplyr::filter(body.part %in% c("Body", "Ovary","Testes")) %>%
   dplyr::filter(Stage == "adult") %>%
   mutate_at("Family", ~replace_na(.,"0")) %>%
  ggplot(aes(y=Ploidy, x=Sex, fill = Sex, lable = Stage)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_classic() +  geom_jitter(width=0.1, size=2) +
  facet_wrap(~body.part ) + 
      theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks = element_blank(),
    panel.border=element_rect(colour="black",size=1, fill = NA))+
ggtitle('Mite ploidy in differnet body parts') +  ylim(0, 3)

p_fam_body = data %>% dplyr::filter(body.part == "Body") %>%
 dplyr::filter(Stage %in% c("Larvae", "Protonymph", "Deuteronymph", "adult")) %>%
  ggplot(aes(y=Ploidy, x=Sex, fill = Sex, lable = Stage)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_classic() +  geom_jitter(width=0.1, size=2) +
facet_wrap(~Stage, nrow = 1 ) + 
  ggtitle('Mite Ploidy in whole body, in differnet developmental stage') +
theme(axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks = element_blank(),
    panel.border=element_rect(colour="black",size=1, fill = NA))+
  theme(legend.position='right')+  ylim(0, 3)
```

### plot histograms
```{r eval=FALSE, include=FALSE}
#hist plot - Ploidy frequency
data %>% dplyr::filter(body.part == "Body") %>%
 dplyr::filter(Stage %in% c("Larvae", "Protonymph", "Deuteronymph", "adult")) %>%
 ggplot(aes(x=Ploidy, color = Sex, fill = Sex)) + 
  geom_histogram(position="dodge") +
 facet_wrap(~Stage, nrow = 1 ) + 
  geom_density(alpha=.2) +
theme_classic2() +
ggtitle("Ploidy count in whole body of mites of differnet stages") 

data %>% dplyr::filter(body.part == "Body") %>%
 ggplot(aes(x=Ploidy, color = Sex, fill = Sex)) + 
  geom_histogram(position="dodge") +
   geom_density(alpha=.2) +
theme_classic2() +
ggtitle("Ploidy count in mites' whole body
Flow-cytometry. n=57") +
  facet_wrap(~stage_mature)

```

### plot by family
```{r}
p_allFam_body_adult = data %>% dplyr::filter(body.part == "Body") %>%
 dplyr::filter(Stage =="adult") %>%
 #   mutate_at("Family", ~replace_na(.,"0")) %>%
  na.omit() %>%
  ggplot(aes(y=Ploidy, x=Family, fill = Sex, lable = Stage)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_bw() +  geom_jitter(width=0.1, size=2) +
#facet_wrap(~Stage ) + 
  ggtitle('Mite Ploidy in whole body, per family') +
   theme(axis.text.x = element_text(angle = 45)) +  theme(legend.position='right')+  ylim(0, 3)

p_fam_body_1_11_27_3 = data %>% dplyr::filter(body.part == "Body") %>%
  dplyr::filter(Family %in% c("1", "11","27","3")) %>%
  ggplot(aes(y=Ploidy, x=Stage_original, fill = Sex, lable = Stage_original)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_bw() +  geom_jitter(width=0.1, size=2) +
  facet_wrap(~Family , nrow = 1) + ggtitle('Mite Ploidy per family') +
   theme(axis.text.x = element_text(angle = 45)) +  theme(legend.position='none')+  ylim(0, 3)

p_fam_body_2_4 = data %>% dplyr::filter(body.part == "Body") %>%
  dplyr::filter(Family %in% c("2","4")) %>%
  ggplot(aes(y=Ploidy, x=Stage_original, fill = Sex, lable = Stage_original)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_bw() +  geom_jitter(width=0.1, size=2) +
  facet_wrap(~Family ) + ggtitle('Mite Ploidy per family') +
   theme(axis.text.x = element_text(angle = 45)) +  theme(legend.position='none')+  ylim(0, 3)

p_fam_body_5 = data %>% dplyr::filter(body.part == "Body") %>%
  dplyr::filter(Family == "5") %>%
  ggplot(aes(y=Ploidy, x=Stage_original, fill = Sex, lable = Stage_original)) + 
 geom_boxplot(outlier.shape = NA, coef=0 ) +  theme_bw() +  geom_jitter(width=0.1, size=2) +
  facet_wrap(~Family ) + ggtitle('Mite Ploidy per family') +
   theme(axis.text.x = element_text(angle = 45)) +  theme(legend.position='none')+  ylim(0, 3)
```

-----

```{r,  echo=FALSE, fig.asp = 0.5, fig.width = 7, out.width="100%",  fig.cap="all developmental stages, males and females, whole body"}
ggplotly(p_fam_body)
```

```{r,  echo=FALSE, fig.asp = 0.5, fig.width = 7, out.width="100%",  fig.cap="all developmental stages, males and females, in different body parts"}
ggplotly(p_family_adults)
```

-----

### Looking at specific families

the Ploidy of the male offspring varied, depends on the family   

for some families ( 1, 11, 27 and 3) the females Ploidy was higher then males.     
while females looks diploid, males look haploid.  
```{r, echo=FALSE,  fig.width = 7, fig.height = 3, out.width="100%"}
#Mite Ploidy per family
ggplotly(p_fam_body_1_11_27_3)
```
-----

for other families ( 2 and 4) the females Ploidy was similar to that of the male offspring.  

```{r, echo=FALSE,  fig.width = 6, fig.height = 3, out.width="60%"}
ggplotly(p_fam_body_2_4)
```
-----

in one family (number 5) the male offspring Ploidy was mixed:   
one males look diploid, and the other two look haploid. 
```{r, echo=FALSE, fig.width = 3, fig.height = 3, out.width="40%"}
#Mite Ploidy per family
ggplotly(p_fam_body_5)
```
